Evaluating machine translation for assimilation via a gap-filling task

### 2015-03-06 2015-eamt-assim/2015-eamt-assim.tex
more stuff
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@59236 72bbbca6-d526-0410-a7d9-f06f51895060

\usepackage{multirow}
\usepackage{natbib}
\usepackage{color}
\usepackage[small,bf]{caption}
\usepackage{graphicx}
\usepackage{microtype} % This saves some space and makes nice typesetting choices

@@

\newcommand{\comment}[1]{\marginpar{\scriptsize\sf \textcolor{blue}{#1}}}

@@

\maketitle
\renewcommand{\baselinestretch}{0.97} % the undetectable 3% squeeze
\begin{abstract}
  This paper provides additional observations on the viability of a
  strategy independently proposed in 2012 and 2013 for evaluation of
 MT for assimilation purposes. The strategy involves
  human evaluators, who are asked to restore keywords (to fill gaps) in
  reference translations. The evaluation method is applied to two language pairs, Basque--Spanish and Tatar--Russian. To reduce the amount of time required to prepare tasks
  and analyse results, an open-source task management system is
  introduced. The evaluation results show that the gap-filling task
  reflects users' understanding of text, and may be used to measure MT
  quality for assimilation purposes.
\end{abstract}

@@

Despite the fact that, as a result of widespread usage of online MT, assimilation (or gisting) is currently the most frequent application of MT (in 2012, daily output of Google Translate matched the yearly output of human translations \citep{google}), few methodologies are established for assimilation evaluation of MT. The methods include post-editing and comparison by bilingual experts \citep{ginesti09},\comment{MLF: should we also cite the original WMT reference cited by Ginestí et al.?} and multiple choice tests \citep{jones07,trosterud12}. These approaches are often costly and prone to subjectivity: see the discussion in \cite{oregan13}. As an alternative, the modification of \emph{cloze}\comment{\emph{cloze} stands for \emph{closure}.} testing \citep{taylor53} was introduced for assimilation evaluation, first by \citet{trosterud12} as a supplementary, and then by \citet{oregan13} as a stand-alone method. Prior to this, cloze tests have been used to evaluate raw MT quality \citep{vanslype79,somers00}. While these authors ask informants to fill gaps in MT output, \citet{trosterud12} and \citet{oregan13} propose to fill gaps in the reference (human) translation. The gap-filling task models how well users comprehend the key points of the text, as it is roughly equivalent with answering questions.\comment{MLF: check if you agree with this} Thus, the method does not directly evaluate the quality of machine-produced text, but rather its usefulness in understanding the meaning of the original text.

The gap-filling method has been successfully used to evaluate the Basque--English Apertium language pair. In this work we extend the evaluation to two more language pairs: Basque--Spanish and Tatar-Russian. The former pair, while not producing the output suitable for post-editing, is a good example of an assimilation MT system. In addition, Basque and Spanish are not mutually understandable, and therefore constitute a good pair for evaluation. For the latter pair, the evaluation served as a quality check in the period of active development during the Google Summer of Code 2014 programme. In addition to evaluating, we explore new aspects of the experiment: the correlation between evaluators' scores, and the influence of the texts' linguistic domain. To facilitate the evaluation, we introduce an automated system which creates task sets from parallel corpora given a range of parameters (number of gaps in a sentence, hint type, gap filler,\comment{MLF: gap filler or gap marker?} \comment{EA: well, initially we had questions with lemmas in place of gaps, and also multiple choice questions where answer options got generated automatically. This is not used in this evaluation, but it's still in the toolkit. Should I remove?} etc.), checks evaluators' answers, and calculates and reports generalized results. This system is integrated into the Appraise MT evaluation platform \citep{federmann12}; the code is open source and is available on GitHub.\footnote{\phantom{\url{https://github.com/Sereni/Appraise}}}\comment{MLF: check if phantom URL is ok}

@@

\begin{table*}
  \begin{tabular}{|l|l|}
     \hline
     \textbf{Ref}   & Ayudas econ\'{o}micas para el tratamiento de toxicoman\'{i}as en comunidades terap\'{e}uticas no concertadas. \\
     \textbf{Task}   & Ayudas econ\'{o}micas para el \{ \} de toxicoman\'{i}as en comunidades terap\'{e}uticas no concertadas. \\
     \textbf{Src} & Komunitate terapeutiko itundu gabeetan toxikomaniak tratatzeko diru-laguntzak ematea. \\
     \textbf{MT}     & Comunidad terap\'{e}utico pactar gabeetan toxikomaniak las-ayudas de dinero para tratar dar. \\
     \hline
  \end{tabular}
  \caption{An example group of sentences showing the gapped sentence and hint types. Reference, MT and task sentences are in Spanish, the source sentence is in Basque.} 
  \label{table:example}
\end{table*}

\begin{table*}
  \begin{tabular}{|l|p{15cm}|}
%@@@  \textbf{Src (Tatar)} & \rus{???? ?????? ???????; ????? 7 ??????????? ??? ??????? ?????, ????? ??? ????? 100 ???? ??????? ???????.}
%  \hline
%  \textbf{MT (Russian)} & \rus{Половина сказать часы; вам 7 тукталыштан поздний опуститься нужный , после его снова 100 метр сказать идти.} \\
     \hline
     \textbf{Ref} & \rus{Примерно полчаса; вам нужно выйти через 7 остановок, потом пройти ещё около 100 метров.} \\\hline
     \textbf{10\%}   & \rus{Примерно полчаса; вам нужно выйти через 7 \{ \}, потом пройти ещё около 100 метров.} \\
     \textbf{20\%}   & \rus{\{ \} полчаса; вам нужно \{ \} через 7 остановок, \{ \} пройти ещё около 100 метров.} \\
     \textbf{30\%} & \rus{Примерно полчаса; вам нужно \{ \} через 7 \{ \}, потом пройти \{ \} около 100 \{ \}.} \\
     \hline
  \end{tabular}
  \caption{Example of different gap percentage settings for a Russian reference sentence.} 
  \label{table:percentage}
\end{table*}
\comment{MLF the gap percentage table 2 is not referenced in the text; also, i have changed it to have the \textbf{Ref} row in the table?}

This section discusses the reasoning behind the gap-filling method and task structure. The gap-filling method 
of evaluating machine translation for assimilation purposes is based on the following hypothesis: a reader's understanding 
of a given text correlates with the number of words they are able to correctly restore in the text. Therefore, the 
base of an assimilation task is a (reference) sentence, where some of the words are blacked out, or removed. The sentence 
is produced by a human (as opposed to machine-translated), and it is in the language known to evaluators, which is also 
the \emph{target language} of the machine translation system.
The additional elements of the task are what we call hints, or extra sentences that help the participant to understand 
the main sentence. There are two types of hints: first, the \emph{source}, which is semantically equivalent to the \emph{reference}, 
also human-produced, but in the source language of the pair. The second type is the \emph{machine-translated} hint, which 
comes from the machine translation of the \emph{source} sentence. Table~\ref{table:example} shows a sample task, and Figure~\ref{figure:screenshot} shows the task in the online evaluation environment.\comment{Use the highest possible resolution for the figure so that aliasing is imperceptible}

\begin{figure*}
  \centering
\includegraphics[width=0.9\textwidth,resolution=144]{appraise-scr}
 \caption{An example set of sentences in the online environment. The task is Russian legal text with 30\% gaps.}
\label{figure:screenshot}
\end{figure*}

@@

For each sentence, the list of candidate keywords is prepared. It is composed of all the words that 
fall into the allowed PoS list. The number of gaps in the sentence is calculated based on 
sentence length and specified gap density. All reference sentences are longer than 10 words. Finally, the required number of keywords is selected 
from the candidate list in such a manner that the gaps are distributed evenly throughout the sentence. We start at a random word in a sentence and check whether it is a keyword candidate. If yes, we remove it, and move \emph{n} words forward, going back to the beginning of sentence if necessary. The step length \emph{n} is the sentence length divided by the desired number of gaps. If the word is not a keyword, or has already been removed, we look at the next word instead. The process is repeated until the designated number of words has been removed, or until there are no more words in the keyword list.\comment{MLF: O'Regan and Forcada did something lots simpler: why not here? It is probably a bit late to change this}

Keyword removal could be one of the most time-consuming steps in task preparation. It normally requires human effort, because we would like to determine the words that contribute the most to understanding the text as opposed to removing random words. \comment{MLF: the procedure described so far seems automatic. What's manual and time consuming?} In our automatic setup, the above procedure is performed by a script integrated into the task generation pipeline. Parts of speech are determined with Apertium morphological analyser. To control for homonymy, we only allow the word into the candidate list if all of its possible part of speech attributions are on the PoS list. For example, if we only allow nouns on the word list, and the word "fly" receives two possible part of speech attributions from the tagger, noun and verb, it is not considered for the candidate list.

@@

The results are presented in this section. Tables \ref{table:res-eus-spa} and \ref{table:res-tat-rus} show
the proportion and standard deviation of correct answers depending on evaluation mode and gap density.\comment{FMT: We should probably integrate the other tables, and refer to them using labels.} The evaluators' correct answer percentage is averaged over the number of evaluators. In addition to the percentage of correct answers we kept a record of time taken to fill the gaps in one sentence. To reduce the noise from participants who were distracted during evaluation, when
calculating times we remove all the results over 6 minutes (the statistical mode is approximately
two minutes). Typical time taken to complete one question varies from under one minute for tasks without hints and few gaps, to approximately two minutes for tasks with more hints and gaps.

@@

\subsection{Annotator agreement}\comment{MLF cannot get these tables to appear numbered later than the other ones!}
After obtaining the results we calculated Krippendorff's alpha~\citep{krippendorff70} measure to represent annotator agreement, shown in Tables~\ref{table:alpha-eus-spa} and~\ref{table:alpha-tat-rus}. 
\begin{table*}
\centering
  \begin{tabular}{|l|r|r|r|r|}
  \hline
 \multirow{2}{*}{\textbf{Density}} & \multicolumn{4}{|c|}{Evaluation mode} \\\cline{2-5}
                                            & \textbf{MT \& Src} & \textbf{MT} & \textbf{Src} & \textbf{No hint} \\
\hline
10\%&0.496&0.517&0.400&0.124\\
20\%&0.714&0.700&0.358&0.275\\
30\%&0.559&0.430&0.406&0.300\\

@@

\begin{table*}
\centering
  \begin{tabular}{|l|r|r|r|r|}
  \hline
 \multirow{2}{*}{\textbf{Density}} & \multicolumn{4}{|c|}{Evaluation mode} \\\cline{2-5}
                                            & \textbf{MT \& Src} & \textbf{MT} & \textbf{Src} & \textbf{No hint} \\
\hline
10\%&0.598&0.459 &0.711 &0.517 \\
20\%&0.740&0.667 &0.473 &0.261 \\
30\%&0.534&0.581 &0.411 &0.412 \\
\hline
\end{tabular}
\caption {Tatar--Russian: Krippendorff Alpha measure} 
\label{table:alpha-tat-rus}
\end{table*}
We selected this measure because of its compatibility with more than two annotators per task and the missing data (not all the gaps were evaluated). To calculate Krippendorff's alpha we used an algorithm implementation by Thomas Grill,\footnote{http://grrrr.org/data/dev/krippendorff\_alpha/} dividing the answers in each gap into two categories: correct and incorrect. The previously obtained synonym lists were taken into account, i.e. if the two answers are different but both correct, they fall into one category. The measure was calculated separately for each hint and percentage combination.

@@

The experiment may easily be repeated for any language pair (provided a parallel corpus) and
any machine translation system. Based on our experience, we would like to suggest the following amendments to the procedure:
\begin{enumerate}
\item As reported in \citep{oregan13}, unless the evaluation is targeted at a specific text domain, it may be beneficial to include a stylistic variety of texts in the initial corpus. Neighboring sentences on the same topic may assist the users in gap-filling tasks;
\item If possible, increase the number of evaluators, or reduce the number of questions per participant. In the above experiments each participant filled from 110 to 187 gaps, divided into small groups. Reducing the amount of work may increase task completion rate;\comment{MLF's crazy idea: recaptchas?}\comment{EA: the participants did not leave the gaps empty, they just did not attempt the next portions of evaluation after finishing some and getting bored. or did I misunderstand the use of captchas? }
\item To account for the adaptation period, provide training tasks before the main evaluations take place.
\end{enumerate}

>>>>

\usepackage{multirow}
\usepackage{natbib}
\usepackage{color}
\usepackage[small,bf]{caption}
\usepackage{graphicx}
%\usepackage{microtype} % This saves some space and makes nice typesetting choices

@@

\newcommand{\comment}[1]{\marginpar{\scriptsize\sf \textcolor{blue}{#1}}}
%\newcommand{\comment}[1]{}

@@

\maketitle
\renewcommand{\baselinestretch}{0.96} % the undetectable 3% squeeze
\begin{abstract}
  This paper provides additional observations on the viability of a
  strategy independently proposed in 2012 and 2013 for evaluation of
 MT for assimilation purposes. The strategy involves
  human evaluators, who are asked to restore keywords (to fill gaps) in
  reference translations. The evaluation method is applied to two language pairs, Basque--Spanish and Tatar--Russian. To reduce the amount of time required to prepare tasks
  and analyse results, an open-source task management system is
  introduced. The evaluation results show that the gap-filling task
  reflects users' understanding of text, and may be used to measure MT
  quality for assimilation purposes.
\end{abstract}

@@

Despite the fact that, as a result of widespread usage of online MT, assimilation (or gisting) is currently the most frequent application of MT (in 2012, daily output of Google Translate matched the yearly output of human translations \citep{google}), few methodologies are established for assimilation evaluation of MT. The methods include post-editing and comparison by bilingual experts \citep{ginesti09},\comment{MLF: should we also cite the original WMT reference cited by Ginestí et al.?} and multiple choice tests \citep{jones07,trosterud12}. These approaches are often costly and prone to subjectivity: see the discussion in \cite{oregan13}. As an alternative, the modification of \emph{cloze}\comment{\emph{cloze} stands for \emph{closure}.} testing \citep{taylor53} was introduced for assimilation evaluation, first by \citet{trosterud12} as a supplementary, and then by \citet{oregan13} as a stand-alone method. Prior to this, cloze tests have been used to evaluate raw MT quality \citep{vanslype79,somers00}. While these authors ask informants to fill gaps in MT output, \citet{trosterud12} and \citet{oregan13} propose to fill gaps in the reference (human) translation. The gap-filling task models how well users comprehend the key points of the text, as it is roughly equivalent with answering questions. Thus, the method does not directly evaluate the quality of machine-produced text, but rather its usefulness in understanding the meaning of the original text.

The gap-filling method has been successfully used to evaluate the Basque--English Apertium language pair. In this work we extend the evaluation to two more language pairs: Basque--Spanish and Tatar-Russian. The former pair, while not producing the output suitable for post-editing, is a good example of an assimilation MT system. In addition, Basque and Spanish are not mutually understandable, and therefore constitute a good pair for evaluation. For the latter pair, the evaluation served as a quality check in the period of active development during the Google Summer of Code 2014 programme. In addition to evaluating, we explore new aspects of the experiment: the correlation between evaluators' scores, and the influence of the texts' linguistic domain. To facilitate the evaluation, we introduce an automated system which creates task sets from parallel corpora given a range of parameters (number of gaps in a sentence, hint type, gap filler,\comment{MLF: gap filler or gap marker?} \comment{EA: well, initially we had questions with lemmas in place of gaps, and also multiple choice questions where answer options got generated automatically. This is not used in this evaluation, but it's still in the toolkit. Should I remove?} etc.), checks evaluators' answers, and calculates and reports generalized results. This system is integrated into the Appraise MT evaluation platform \citep{federmann12}; the code is open source and is available on GitHub.\footnote{\phantom{\url{https://github.com/Sereni/Appraise}}}

@@

\begin{table*}
  \centering
  \begin{tabular}{|l|l|}
     \hline
     \textbf{Ref}  & Ayudas econ\'{o}micas para el tratamiento de toxicoman\'{i}as en comunidades terap\'{e}uticas no concertadas. \\
     \textbf{Task} & Ayudas econ\'{o}micas para el \{ \} de toxicoman\'{i}as en comunidades terap\'{e}uticas no concertadas. \\
     \textbf{Src}  & Komunitate terapeutiko itundu gabeetan toxikomaniak tratatzeko diru-laguntzak ematea. \\
     \textbf{MT}   & Comunidad terap\'{e}utico pactar gabeetan toxikomaniak las-ayudas de dinero para tratar dar. \\
     \hline
  \end{tabular}
  \caption{An example group of sentences showing the gapped sentence and hint types. Reference, MT and task sentences are in Spanish, the source sentence is in Basque.} 
  \label{table:example}
\end{table*}

\begin{table*}
  \centering
  \begin{tabular}{|l|l|}
     \hline
     \textbf{Ref} & \rus{{\small Примерно полчаса; вам нужно выйти через 7 остановок, потом пройти ещё около 100 метров.}} \\
     \hline
     \textbf{10\%} & \rus{{\small Примерно полчаса; вам нужно выйти через 7 \{ \}, потом пройти ещё около 100 метров.}} \\
     \textbf{20\%} & \rus{{\small \{ \} полчаса; вам нужно \{ \} через 7 остановок, \{ \} пройти ещё около 100 метров.}} \\
     \textbf{30\%} & \rus{{\small Примерно полчаса; вам нужно \{ \} через 7 \{ \}, потом пройти \{ \} около 100 \{ \}.}} \\
     \hline
  \end{tabular}
  \caption{Example of different gap percentage settings for a Russian reference sentence.} 
  \label{table:percentage}
\end{table*}
\comment{MLF the gap percentage table 2 is not referenced in the text; also, i have changed it to have the \textbf{Ref} row in the table?}

This section discusses the reasoning behind the gap-filling method and task structure. The gap-filling method 
of evaluating machine translation for assimilation purposes is based on the following hypothesis: a reader's understanding 
of a given text correlates with the number of words they are able to correctly restore in the text. Therefore, the 
base of an assimilation task is a (reference) sentence, where some of the words are blacked out, or removed. The sentence 
is produced by a human (as opposed to machine-translated), and it is in the language known to evaluators, which is also 
the \emph{target language} of the machine translation system.
The additional elements of the task are what we call hints, or extra sentences that help the participant to understand 
the main sentence. There are two types of hints: first, the \emph{source}, which is semantically equivalent to the \emph{reference}, 
also human-produced, but in the source language of the pair. The second type is the \emph{machine-translated} hint, which 
comes from the machine translation of the \emph{source} sentence. Table~\ref{table:example} shows a sample task, and Figure~\ref{figure:screenshot} shows the task in the online evaluation environment.
%\comment{Use the highest possible resolution for the figure so that aliasing is imperceptible}

\begin{figure*}
  \centering
\includegraphics[width=1.0\textwidth,resolution=144]{appraise-scr}
 \caption{An example set of sentences in the online environment. The task is Russian legal text with 30\% gaps.}
\label{figure:screenshot}
\end{figure*}

@@

For each sentence, the list of candidate keywords is prepared. It is composed of all the words that 
fall into the allowed PoS list. The number of gaps in the sentence is calculated based on 
sentence length and specified gap density. All reference sentences are longer than 10 words. Finally, the required number of keywords is selected 
from the candidate list in such a manner that the gaps are distributed evenly throughout the sentence. We start at a random word in a sentence and check whether it is a keyword candidate. If yes, we remove it, and move \emph{n} words forward, going back to the beginning of sentence if necessary. The step length \emph{n} is the sentence length divided by the desired number of gaps. If the word is not a keyword, or has already been removed, we look at the next word instead. The process is repeated until the designated number of words has been removed, or until there are no more words in the keyword list.
%\comment{MLF: O'Regan and Forcada did something lots simpler: why not here? It is probably a bit late to change this}

%\comment{MLF: the procedure described so far seems automatic. What's manual and time consuming?} 
Keyword removal could be one of the most time-consuming steps in task preparation. It normally requires human effort, because we would like to determine the words that contribute the most to understanding the text as opposed to removing random words. In our automatic setup, the above procedure is performed by a script integrated into the task generation pipeline. Parts of speech are determined with Apertium morphological analyser. To control for homonymy, we only allow the word into the candidate list if all of its possible part of speech attributions are on the PoS list. For example, if we only allow nouns on the word list, and the word "fly" receives two possible part of speech attributions from the tagger, noun and verb, it is not considered for the candidate list.

@@

%\comment{FMT: We should probably integrate the other tables, and refer to them using labels.} 
The results are presented in this section. Tables \ref{table:res-eus-spa} and \ref{table:res-tat-rus} show
the proportion and standard deviation of correct answers depending on evaluation mode and gap density.
The evaluators' correct answer percentage is averaged over the number of evaluators. In addition to the percentage of correct answers we kept a record of time taken to fill the gaps in one sentence. To reduce the noise from participants who were distracted during evaluation, when
calculating times we remove all the results over 6 minutes (the statistical mode is approximately
two minutes). Typical time taken to complete one question varies from under one minute for tasks without hints and few gaps, to approximately two minutes for tasks with more hints and gaps.

@@

\subsection{Annotator agreement}
After obtaining the results we calculated Krippendorff's alpha~\citep{krippendorff70} measure to represent annotator agreement, shown in Tables~\ref{table:alpha-eus-spa} and~\ref{table:alpha-tat-rus}. 
\begin{table*}
\centering
  \begin{tabular}{|l|r|r|r|r|}
  \hline
 \multirow{2}{*}{\textbf{Density}} & \multicolumn{4}{|c|}{Evaluation mode} \\\cline{2-5}
                                            & \textbf{MT \& Src} & \textbf{MT} & \textbf{Src} & \textbf{No hint} \\
\hline
10\%&0.496&0.517&0.400&0.124\\
20\%&0.714&0.700&0.358&0.275\\
30\%&0.559&0.430&0.406&0.300\\

@@

\begin{table*}
\centering
  \begin{tabular}{|l|r|r|r|r|}
  \hline
 \multirow{2}{*}{\textbf{Density}} & \multicolumn{4}{|c|}{Evaluation mode} \\\cline{2-5}
                                            & \textbf{MT \& Src} & \textbf{MT} & \textbf{Src} & \textbf{No hint} \\
\hline
10\%&0.598&0.459 &0.711 &0.517 \\
20\%&0.740&0.667 &0.473 &0.261 \\
30\%&0.534&0.581 &0.411 &0.412 \\
\hline
\end{tabular}
\caption {Tatar--Russian: Krippendorff Alpha measure} 
\label{table:alpha-tat-rus}
\end{table*}
We selected this measure because of its compatibility with more than two annotators per task and the missing data (not all the gaps were evaluated). To calculate Krippendorff's alpha we used an algorithm implementation by Thomas Grill,\footnote{\url{http://grrrr.org/data/dev/krippendorff_alpha/}} dividing the answers in each gap into two categories: correct and incorrect. The previously obtained synonym lists were taken into account, i.e. if the two answers are different but both correct, they fall into one category. The measure was calculated separately for each hint and percentage combination.

@@

The experiment may easily be repeated for any language pair (provided a parallel corpus) and
any machine translation system. Based on our experience, we would like to suggest the following amendments to the procedure:
\begin{enumerate}
\item As reported in \cite{oregan13}, unless the evaluation is targeted at a specific text domain, it may be beneficial to include a stylistic variety of texts in the initial corpus. Neighboring sentences on the same topic may assist the users in gap-filling tasks;
\item If possible, increase the number of evaluators, or reduce the number of questions per participant. In the above experiments each participant filled from 110 to 187 gaps, divided into small groups. Reducing the amount of work may increase task completion rate;\comment{MLF's crazy idea: recaptchas?}\comment{EA: the participants did not leave the gaps empty, they just did not attempt the next portions of evaluation after finishing some and getting bored. or did I misunderstand the use of captchas? }
\item To account for the adaptation period, provide training tasks before the main evaluations take place.
\end{enumerate}
